---
title: "BioDyn Summary Document"
author: "Shyamolina"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: yes
    number_sections: no
  pdf_document:
    toc: yes
urlcolor: blue
---

```{r setup, include=FALSE, echo=F, message=F, results="hide"}
knitr::opts_chunk$set(echo = TRUE)
options(scipen = 1, digits = 4)
seed<-101
library(tidyverse)
library(knitr)
library(kableExtra)
```

# Objective\label{Objective}
  - to assess community stability considering synchronous interactions between species at their extreme abundance (common or rare) for both realms: freshwater and terrestrial
  - usual method uses species richness (R) and variance ratio (VR) to assess stability with basic expectation: stability should increase with richness and decrease with synchrony. VR is a way to express the level of synchrony between species. VR=1 (independent interaction, zero correlation between species), VR<1 (asynchonous interaction, -ve correlation between species), VR>1 (synchronous interaction, +ve correlation between species).
  - we want to see if we add additional predictors, e.g., (Asymmetry = A, direction of asymmetry = uniA, skewness ratio = SR), then can we explain the stability better?

# Overview of Data\label{OverData}

We collected species-level community data from freshwater and terrestrial realms. Here is a snippet of the cleaned metadata with 28 observations:

```{r metadataview, echo=F}

df_md<-read.csv("./Results/gather_res/metadata_summary.csv")
df_md$UID<-paste(df_md$source,df_md$STUDY_ID,sep=",")
df_md<-df_md%>%select(UID,newsite,REALM,TAXA,ORGANISMS,CENT_LAT,CENT_LONG,nsp,nyr)
df_md_r<-df_md[c(1:3,823:827,859:862,2605:2606,2642:2645,2766:2768,2773,2774,2788,2789,2790,2862:2863),]
rownames(df_md_r)<-NULL
df_md_r %>%
  kbl() %>%
  kable_styling()
```

Data (total = `r nrow(df_md)` observations) have the following hierarchical structure: 

 - Realms: 2 levels
```{r metadataview2, echo=F}
table(df_md$REALM)
```
 - Taxa: total 7 levels nested within Realms
```{r metadataview3, echo=F}
df_md%>%group_by(REALM)%>%count(TAXA)%>%ungroup()
```
 - UID(= combination of Datasource and STUDY_ID): total 249 levels nested within TAXA, each    level might have 1 or multiple observations of **newsite**
```{r metadataview4, echo=F}
df_md%>%group_by(REALM, TAXA)%>%count(UID)%>%ungroup()
```

# Model\label{Model}

Fig.1 nice caption
```{r, out.height = "460px", out.width='800px', echo=F}
knitr::include_graphics("./Results/gather_res/f_nL_f_nU_ratio_for_eachrealm.pdf")
```


I think \ref{OverData} is good.


