---
title: A global analysis of community stability across freshwater and terrestrial
  realms
author: "Shyamolina"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    number_sections: no
    keep_tex: yes
    fig_caption: yes
  word_document: default
header-includes:
- \usepackage{setspace}\doublespacing
- \usepackage{lettrine}
- \usepackage{hyperref}
- \usepackage{lineno}\linenumbers
- \hypersetup{linkcolor=red}
- \usepackage{caption}\captionsetup[figure]{width=16cm}\captionsetup[figure]{font={stretch=1.5}}\captionsetup[table]{width=16cm}\captionsetup[table]{font={stretch=1.5}}
- \usepackage[nomarkers,nolists]{endfloat}
mainfont: Times New Roman
tables: yes
link-citations: yes
urlcolor: blue
indent: yes
---

```{r setup, include=FALSE, echo=F, message=F, results="hide"}
knitr::opts_chunk$set(echo = TRUE)
options(scipen = 1, digits = 4)
seed<-101
library(tidyverse)
library(knitr)
library(kableExtra)
```

<!--
Title1: Community stability and asymmetric species interactions differ across realms: freshwater vs. terrestrial
Title 2: Role of asymmetric species interactions in community stability: a comparative study between freshwater and terrestrial realm
Title 3: Asymmetric interactions and diversity-synchrony-stability relationship: a comparative study between freshwater and terrestrial realm
Title 4: Asymmetric interactions and diversity-synchrony-stability relationship differs in freshwater vs. terrestrial realm


Authors: Shyamolina and others

Abstract (200 w)/ Body (max 2500 w including abstract)/ Display items (max 4) for a standard 5-page article - Nature
Research Articles (up to ~4500 words, including references, notes and captions–corresponds to ~5 printed pages in the journal) are expected to present a major advance. Research Articles include an abstract, an introduction, up to 6 figures or tables, sections with brief subheadings, and about 40 references. Materials and Methods should be included in supplementary materials, which should also include information needed to support the paper's conclusions. - Science
TREE, opinion - 2500 w
Abstract (150 w)/ Body (3500 w)/ Display items (max 6) - Nature eco & evo
Abstract (250 w)/ sig statement (50-120 w)/ A standard 6-page article is approximately 4,000 words, 50 references, and 4 medium-size graphical elements (i.e., figures and tables). - PNAS

-->

# Abstract\label{Abstract}

Research on stability in ecological communities aims to understand different drivers for stability under the ongoing global changes, however, there is no comparative study of how stability-drivers relationships and interspecific interactions at the extremes vary across the realm (i.e. terrestrial vs. aquatic). Such a study is important to understand how and possibly why are the realms different in stability and how drivers have different impacts in different realms. We address this research gap using a hierarchical mixed-effect Bayesian model for an extensive dataset of 2759 communities (716 freshwater and 2043 terrestrial study sites across 7 taxa). Stability is supposed to increase with increasing species richness (portfolio effect) and with increasing asynchrony (insurance hypothesis). Variance ratio, based on covariance measure, quantifies the degree of (a)synchrony in a community with respect to random fluctuation between species. We proposed an additional predictor for stability - total tail asymmetry - that sums the difference in the pairwise associations of species' individual ranks throughout the study periods at either tail of their joint distribution, i.e., association in low rank-low rank vs. high rank-high rank. Total tail asymmetry quantifies asymmetric patterns of synchrony on a community level, i.e. if species are overall synchronously rare (or common) through a stronger association between low (or high) ranks of species pair. We showed: (i) stability is different across realms - freshwaters are less stable than terrestrial ones; (ii) drivers have different impacts on stability based on the realm (freshwater vs. terrestrial); and (iii) including total tail asymmetry (i.e. asymmetry in pairwise species synchrony) in models have a higher performance score than just using two classic predictors - richness and variance ratio. Linking synchrony patterns with stability in multiple realms offers new insights into the mechanistic pathway to explain variation in community stability.


# Main Body\label{Main_Body}
    
The concept of community stability is central to our understanding of how biodiversity is maintained in ecological communities (Loreau and de Mazancourt 2013; McCann 2000; Tilman, Lehman, and Bristow 1998). Empirical ecologists, in general, use the inverse of variability to assess community stability throughout the study period - that means a community with less variation in overall productivity through time is considered to be more stable. Community stability increases with species richness or diversity, i.e., the portfolio effect (Schindler, Armstrong, and Reed 2015; Doak et al. 1998), and when fluctuations between species abundances over time tend to be asynchronous, i.e., the insurance hypothesis (Naeem and Li 1997; Yachi and Loreau 1999). This emphasizes the importance of specific types of species interactions for community stability (Mougi and Kondoh 2012; Lehman and Tilman 2000). Synchrony (or asynchrony) in communities is classically inferred from a variance ratio (VR) approach that compares the sum of temporal covariance between all species pairs present in the community with respect to its baseline variance, i. e., of an independent community. As all species in an independent community are believed to fluctuate randomly, the covariance between any dissimilar species pair would vanish. Therefore, a community having higher (or lower) VR than its independent analog means species are fluctuating synchronously (or asynchronously) with time in that community. An alternative approach to compute variance ratio also used a community with perfect synchrony among all species as the baseline (Loreau and de Mazancourt 2008). However, the above-mentioned measures of synchrony via variance ratio approach fail to capture any association between species pairs at their extremes, particularly, the dependence in the tail of species joint distributions.

# Stability and the patterns of synchrony

Because the diversity-stability relationship is influenced by climatic extremes (Boeck et al. 2018) and extreme events could change the interspecific interactions at the tails in substantial ways to affect community stability (Ghosh, Cottingham, and Reuman 2021), we focus on the finer scale resolution of synchrony to interpret variation in stability. In other words, we ask when does the synchrony happen in the community - is it when species pairs are synchronously rare, or is it when species pairs are synchronously abundant. We measured the tail-dependence between time series for a given species pair where each species timeseries was ranked individually and normalized (for details see Methods). The three common patterns in interspecific interactions are independent, synchrony, and asynchrony based on overall correlations - either zero, positive or negative, respectively (Fig. 1a), however, on a finer scale resolution, an asymmetric pattern with lower (or upper) tail-dependence arises when two synchronous species have simultaneous low (or high) ranks for the same study years (Figs. 1b-c). We repeated the tail dependence analysis (Ghosh, Sheppard, Reid, et al. 2020; Ghosh, Sheppard, Holder, et al. 2020) for all possible dissimilar species pairs to get a general pattern of how much tail asymmetry was there for a given community (Fig. 1d). We hypothesize that the total tail asymmetry (TA), the sum of tail asymmetry in either direction, would decrease community stability because species interactions at their extremes are generally driven by extreme environmental perturbations.

# Drivers that influence community stability between freshwater and terrestrial realms

To test if patterns of synchrony with tail asymmetry could explain the variation in community stability better rather than just using classic predictors (e.g., richness and variance ratio), we took four alternative models as the building block (Fig. 1e-h). We started with richness, R, as the single predictor for stability (Fig. 1e), then considered a model with two predictors - richness (R) and synchrony (variance ratio, VR) (Fig. 1f). The other two models considered the asymmetric pattern of synchrony as one of the predictors for stability (Fig. 1g-h). 
Because the ongoing climate change has been found to affect the stability in both terrestrial and aquatic realms (Ma et al. 2017; Boeck et al. 2018; Antão et al. 2020), we also wanted to test our models including the realm-effect. Therefore, we collated abundance (or biomass) timeseries data for each species present in a plot nested within a study site (hence each plot is considered as a community) for both terrestrial and freshwater realms. Most of the data were identified on the species level, when species-level identification was not available/ confident throughout the sampling periods (mainly for freshwater insects, planktons), higher taxonomic level aggregation was performed. We maintained consistent sampling efforts during data filtering (e.g. considering nearly the same sets of sampling months to compute the annual average). We applied two inclusion criteria - (i) a minimum duration for each study should be 20 years, and (ii) species that were present for at least 70$\%$ of the study period were included in our analysis - this is richness, R, used in our model. That resulted in a dataset of 2759 communities (i.e. plots) across 7 taxa nested within 244 study sites (716 freshwater and 2043 terrestrial communities across 7 taxa - fish: 588, freshwater invertebrates: 113, freshwater plants: 15, birds: 1261, terrestrial invertebrates: 159, terrestrial plants: 440, terrestrial mammals: 183). However, there was an imbalance between realms (terrestrial:freshwater ~ 74:26) because of the huge amount of bird data from the BBS database (~1227 birds data). Therefore, we reduced that bias for birds using stratified random sampling (see Methods) so that to get a more balanced dataset (terrestrial:freshwater ~ 60:40, birds: 293) as shown in Fig. A1. 
With this reduced dataset, we built a hierarchical mixed-effect Bayesian model set up, where predictors were considered as the fixed effects, and random effects were due to plots nested within the study sites for different taxa. The model tested the performance for each one of the four models (Fig. 1e-h) without and with the realm-effect included to explain the variation in the response variable, i.e., the temporal stability (see Methods for details). Stability was computed by taking the ratio of median and the interquartile range of distribution for the total community-abundance (or biomass) timeseries.

# Asymmetry in synchrony matters for community stability 

When we did not include the realm, the model showed a positive coefficient for richness and a negative coefficient for synchrony (with or without tail asymmetry) in explaining the variation in community stability (Fig. 2a-d). The effects were significant and consistent as we expected from Fig. 1e-h. Including asymmetric patterns in synchrony had higher marginal R2 (i.e. variation explained due to fixed effect) than the models with only the classic predictors - richness and/ or variance ratio (Table 1, column 8 and rows 5-8). Model performance score, calculated based on the common statistics summarized in columns in Table 1, also indicated that the total tail asymmetry if included as one of the predictors could outperform the other models that had considered classic predictors only.
	The above-mentioned feature remained consistent when we included the effect of realms in the model, i.e., models with asymmetric pattern in synchrony performed better in explaining the variation for stability in freshwater and terrestrial realm than the models with classic predictors in those realms (rows 1-4 in Table 1). Our result showed asymmetric species interactions have a significant role in explaining community stability, and in the future, due to frequent extreme events in climate, this asymmetric pattern could be more prominent for sustaining such communities across the realm.

# Realm matters: drivers have different impacts on stability in freshwater and terrestrial communities

We found stability also varied across the realm, freshwater communities are less stable than the terrestrial ones (compare the intercepts, Fig. 2 e-l), though the difference was moderately significant (Fig. 3a). Moreover, adding realm as a fixed effect interacting with the predictors (e.g. richness, variance ratio, or total tail asymmetry) increased the model performance (compare rows 1 and 5; 2 and 6; 3 and 7; 4 and 8 from Table 1). This explained one should not ignore the realm when considering long-term community dynamics. The predictors also affect stability at different strengths in the freshwater and terrestrial realm (Fig. 2e-l). Species richness, variance ratio, and total tail asymmetry were strongly significant in assessing the variation for stability in the terrestrial realm (Fig. 2i-l), but in the freshwater realm, the effects were not always significant (Fig. 2e-h). While variance ratio and richness had a strong to moderate effect on stability, total tail asymmetry had a weak effect on stability for the freshwater realm. Though the effect of different predictors may or may not be significant in each realm individually, the basic difference between their effect from both realms was always significant (Fig. 3b-d). Moreover, if we focus on the effect size for different taxa, freshwater communities showed a consistent and significant negative trend, whereas for terrestrial communities there was no significant trend.  

Discussion 

While many studies (e.g. (Valencia et al. 2020; Aoki 2003) have considered relationships between diversity, stability, and synchrony within single realms, our study is one of the first to compare the stability-drivers patterns across the terrestrial and freshwater realms including asymmetric interactions between species. Considering the difference between realms is important to prioritize conservation strategies for specific communities as often found freshwater systems’ needs are different than terrestrial ones (Leal et al. 2020). Also, it could help prioritize the target area, where both realms’ priority overlap, e.g., could be based on decreasing stability - one of the recommendations used in systematic conservation planning (Amis et al. 2009). 

Our study showed freshwater communities are less stable than terrestrial ones, and also the drivers had more impact on terrestrial communities. That was expected due to gradients of richness across realms (May et al. 1994; Grosberg, Vermeij, and Wainwright 2012). As terrestrial realms are more species-rich than freshwater ones, it would make a terrestrial community more stable (or less variable) due to the positive portfolio effect. Also, more species in a community had a higher chance to be asynchronous which in turn would provide additional stability in terrestrial realms (insurance hypothesis). For example, a study showed temporal stability in forests increased as asynchrony increased with increasing richness (Morin et al. 2014). Fig. A3 in the appendix showed freshwater realm had more synchrony than asynchrony - so expected to be less stable. 
We also considered asymmetric species interactions for community stability - tail asymmetry in synchrony is common in ecological response (Ghosh, Sheppard, Holder, et al. 2020) and has prominent effects on population extinction risk (Ghosh, Sheppard, and Reuman 2020) and community stability (Ghosh, Cottingham, and Reuman 2021). Asymmetric species-interactions were strongly significant in explaining variation for terrestrial community stability, but in freshwater communities, they had weak effects (Fig. 2g,k; h,l). Both realms showed an expected relationship as in Fig. 1g-h: total tail asymmetry destabilized the community. The intuition was: with increasing asymmetry, stability should decrease as the community experienced frequent extreme events, species were becoming simultaneously rare or common (more outliers in the interactions) compared to a situation if there were no such disturbance. Terrestrial realms have more heterogeneous environments than freshwater realms, so more extreme events (or disturbance) would be expected in terrestrial communities leading to more asymmetric interactions. We have seen temperature affects the realm differently (Antão et al. 2020). Tail asymmetry (or extremes) was not that effective in freshwater, probably because drastic changes in temperature are better buffered in water than land, reducing the likelihood of local extinction, and also terrestrial realm has more microhabitats (or refugia) for rescue effect. A future study incorporating different environmental variables and their extremes in the model could be useful to see the direct and indirect effects on community stability. 

We found the communities were indeed different based on the realm, the difference between taxa was not prominent that much because of varying sample sizes. To detect the stability across taxa, one could use a sample-based rarefaction method using taxa as a fixed effect in the model. Another important thing to consider is the temporal variation or generation time for each taxon. Though in our model, we considered random effect due to taxa as a proxy of generation time as our main focus was on realm - terrestrial vs. freshwater. In our study, we included species that present at least 70$\%$ of the sampling periods, as we found some support that it was the dominant species that were more important for the stability of the community (Sasaki and Lauenroth 2011; Valencia et al. 2020). Our study first showed the difference between terrestrial and freshwater communities from a stability-drivers mechanism perspective using a huge long-term (min. 20 years) dataset and considering the interspecific interactions at extremes, in particular. As a recent study (Maxwell et al. 2019) discussed the importance to consider the full range and severity of ecological responses to extreme events for improving conservation efforts, we believe, our study will help to initiate conservation management across realms by considering heterogeneity in species synchrony pattern to extreme disturbances.

# DATA AVAILABILITY
Mostly our data are a compilation of several open-access databases (e.g. BioTIME, BBS, RivFishTIME), few are private data and can be made available upon request, metadata is available as an appendix.

# CODE AVAILABILITY
Complete computer code written in R programming language for this project is at https://github.com/sghosh89/BioDyn.

# Competing interests
We declare we have no competing interests.

# FUNDING
We thank the ETH Board for funding through the Blue-Green Biodiversity (BGB) Initiative 2020.

# METHOD SUMMARY

<!--
\textbf{Quantifying total tail-asymmetry:} We started with the abundance (or biomass) time series of Nspecies in the community that was sampled for T (20) years. We then measured the Spearman correlation for each possible pairwise combination of species time series (total NC2 combinations possible) and found different patterns (independent: no correlation, synchronous: positive correlation, asynchronous: negative correlation). To measure tail-asymmetry in synchronous patterns, we followed these steps: (i) we made a ‘copula plot’  following (Ghosh, Sheppard, Holder, et al. 2020) with each pair - where each time series was ranked independently and divided by (T+1) so that they fell within a unit box (0,1), these we called “normalized rank plot”. You can see how the rank of one species relates to the rank of the other species. When ranks of both species matched, they were closely associated and synchrony happened. If there was more association between lower ranks (i.e. when both species were rare as in fig. 1b) - a strong tail appeared in the lower half of the diagonal line y= -x+1line. Conversely, if there was more association between higher ranks (i.e. when both species were common as in Fig. 1c) - a strong tail appeared in the upper half of the diagonal line y= -x+1line. (ii) we then computed partial Spearman correlation (Ghosh, Sheppard, Holder, et al. 2020) for the lower half (Corl) and upper half (Coru) of the copula plot. For plots like Fig. 1b, Corl > Coru, i.e., the data points below the diagonal line y= -x+1of the normalized rank plot contributes more to the total Spearman correlation of all the data points. Similarly, for plots like Fig. 1c, Corl < Coru. (iii) We then quantified total tail asymmetry as, TA =∑|Corl - Coru|ij, where the sum was for any two dissimilar species i and j such that i < j, i = 1,...., N, j=1,..., N. This is the measure of tail asymmetry from either tail (lower or upper) for all possible dissimilar species pairs from the community.  

\textbf{Technical details of hierarchical Bayesian modeling:} Bayesian analyses were performed in R-software using the ‘brm’ function from the 'brms' package after data standardization (zero mean, unit standard deviation) to compare coefficients for different types of predictors on the same scale. Parameters used for this study: 16000 total iterations with 75$\%$ warmup on each of four chains (starting with randomized initialization) to fit the parameters of a gaussian distribution with weakly informative priors, where every fourth sample was drawn using NUTS sampling from post-warm up samples to avoid auto-correlation. Convergence was assured by checking Rhat = 1 (potential scale reduction factor on split chains) and significance was assessed on a moderate (75$\%$ CI) to strong (95$\%$ CI) scale. Comparison table among model performance was computed using compare_performance() function from the 'performance' R-package.

\textbf{Stratified random sampling for birds from the BBS database:} Initially, there was huge birds data (n=1227, min. 20 years sampled) from the BBS database across 70 strata from North America and Canada. We reduced this dataset nearly to its 21$\%$ by randomly sampling data points from each stratum proportional to the strata-area, i.e., more sampling sites would be expected from a bigger stratum than a smaller stratum. This sampling resulted in 259 data points from BBS bird data sampled across 70 strata, which appeared close to the average over all other terrestrial taxa (~260). Code for this sampling technique is given in [GitHub](https://github.com/sghosh89/BioDyn/blob/main/R/gather_res/subset_BBS_21.R).
-->


<!--References
Amis, Mao A., Mathieu Rouget, Mervyn Lotter, and Jenny Day. 2009. “Integrating Freshwater and Terrestrial Priorities in Conservation Planning.” Biological Conservation 142 (10): 2217–26.
Antão, Laura H., Amanda E. Bates, Shane A. Blowes, Conor Waldock, Sarah R. Supp, Anne E. Magurran, Maria Dornelas, and Aafke M. Schipper. 2020. “Temperature-Related Biodiversity Change across Temperate Marine and Terrestrial Systems.” Nature Ecology & Evolution 4 (7): 927–33.
Aoki, Ichiro. 2003. “Diversity-Productivity-Stability Relationship in Freshwater Ecosystems: Whole-Systemic View of All Trophic Levels.” Ecological Research. https://doi.org/10.1046/j.1440-1703.2003.00564.x.
Boeck, Hans J. De, Hans J. De Boeck, Juliette M. G. Bloor, Juergen Kreyling, Johannes C. G. Ransijn, Ivan Nijs, Anke Jentsch, and Michaela Zeiter. 2018. “Patterns and Drivers of Biodiversity-Stability Relationships under Climate Extremes.” Journal of Ecology. https://doi.org/10.1111/1365-2745.12897.
Doak, D. F., D. Bigger, E. K. Harding, M. A. Marvier, R. E. O’Malley, and D. Thomson. 1998. “The Statistical Inevitability of Stability-Diversity Relationships in Community Ecology.” The American Naturalist 151 (3): 264–76.
Ghosh, Shyamolina, Kathryn L. Cottingham, and Daniel C. Reuman. 2021. “Species Relationships in the Extremes and Their Influence on Community Stability.” Philosophical Transactions of the Royal Society of London. Series B, Biological Sciences 376 (1835): 20200343.
Ghosh, Shyamolina, Lawrence W. Sheppard, Mark T. Holder, Terrance D. Loecke, Philip C. Reid, James D. Bever, and Daniel C. Reuman. 2020. “Copulas and Their Potential for Ecology.” In Advances in Ecological Research, 62:409–68. Elsevier.
Ghosh, Shyamolina, Lawrence W. Sheppard, Philip C. Reid, and Daniel Reuman. 2020. “A New Approach to Interspecific Synchrony in Population Ecology Using Tail Association.” Ecology and Evolution 10 (23): 12764–76.
Ghosh, Shyamolina, Lawrence W. Sheppard, and Daniel C. Reuman. 2020. “Tail Associations in Ecological Variables and Their Impact on Extinction Risk.” Ecosphere. https://doi.org/10.1002/ecs2.3132.
Grosberg, Richard K., Geerat J. Vermeij, and Peter C. Wainwright. 2012. “Biodiversity in Water and on Land.” Current Biology: CB 22 (21): R900–903.
Leal, Cecília G., Gareth D. Lennox, Silvio F. B. Ferraz, Joice Ferreira, Toby A. Gardner, James R. Thomson, Erika Berenguer, et al. 2020. “Integrated Terrestrial-Freshwater Planning Doubles Conservation of Tropical Aquatic Species.” Science 370 (6512): 117–21.
Lehman, Clarence L., and David Tilman. 2000. “Biodiversity, Stability, and Productivity in Competitive Communities.” The American Naturalist 156 (5): 534–52.
Loreau, Michel, and Claire de Mazancourt. 2008. “Species Synchrony and Its Drivers: Neutral and Nonneutral Community Dynamics in Fluctuating Environments.” The American Naturalist 172 (2): E48–66.
———. 2013. “Biodiversity and Ecosystem Stability: A Synthesis of Underlying Mechanisms.” Ecology Letters 16 Suppl 1 (May): 106–15.
Maxwell, Sean L., Nathalie Butt, Martine Maron, Clive A. McAlpine, Sarah Chapman, Ailish Ullmann, Dan B. Segan, and James E. M. Watson. 2019. “Conservation Implications of Ecological Responses to Extreme Weather and Climate Events.” Diversity & Distributions 25 (4): 613–25.
May, Robert Mccredie, J. Godfrey, John Rex Beddington, David Henry Cushing, Robert Mccredie May, and John Hyslop Steele. 1994. “Biological Diversity: Differences between Land and Sea.” Philosophical Transactions of the Royal Society of London. Series B, Biological Sciences 343 (1303): 105–11.
Ma, Zhiyuan, Huiying Liu, Zhaorong Mi, Zhenhua Zhang, Yonghui Wang, Wei Xu, Lin Jiang, and Jin-Sheng He. 2017. “Climate Warming Reduces the Temporal Stability of Plant Community Biomass Production.” Nature Communications. https://doi.org/10.1038/ncomms15378.
McCann, Kevin Shear. 2000. “The Diversity–stability Debate.” Nature 405 (6783): 228–33.
Morin, Xavier, Lorenz Fahse, Claire de Mazancourt, Michael Scherer-Lorenzen, and Harald Bugmann. 2014. “Temporal Stability in Forest Productivity Increases with Tree Diversity due to Asynchrony in Species Dynamics.” Ecology Letters 17 (12): 1526–35.
Mougi, A., and M. Kondoh. 2012. “Diversity of Interaction Types and Ecological Community Stability.” Science 337 (6092): 349–51.
Naeem, Shahid, and Shibin Li. 1997. “Biodiversity Enhances Ecosystem Reliability.” Nature 390 (6659): 507–9.
Sasaki, Takehiro, and William K. Lauenroth. 2011. “Dominant Species, rather than Diversity, Regulates Temporal Stability of Plant Communities.” Oecologia 166 (3): 761–68.
Schindler, Daniel E., Jonathan B. Armstrong, and Thomas E. Reed. 2015. “The Portfolio Concept in Ecology and Evolution.” Frontiers in Ecology and the Environment 13 (5): 257–63.
Tilman, D., C. L. Lehman, and C. E. Bristow. 1998. “Diversity-Stability Relationships: Statistical Inevitability or Ecological Consequence?” The American Naturalist 151 (3): 277–82.
Valencia, Enrique, Francesco de Bello, Thomas Galland, Peter B. Adler, Jan Lepš, Anna E-Vojtkó, Roel van Klink, et al. 2020. “Synchrony Matters More than Species Richness in Plant Community Stability at a Global Scale.” Proceedings of the National Academy of Sciences of the United States of America 117 (39): 24345–51.
Yachi, S., and M. Loreau. 1999. “Biodiversity and Ecosystem Productivity in a Fluctuating Environment: The Insurance Hypothesis.” Proceedings of the National Academy of Sciences of the United States of America 96 (4): 1463–68.-->




```{r, out.height = "500px", out.width='500px', echo=F}
# example fig for html output
#knitr::include_graphics("./conceptual_fig_cop_ai_af.pdf")
```

\begin{figure}[!ht]
\includegraphics[width=18cm]{./Results/gather_res/simple_subset_birds_21/introfig_simple_ai2.pdf}
\caption{conceptual figure (a): three common patterns for pairwise species interactions: independent, synchrony, asynchrony; (b-c) synchrony patterns with tail dependence: synchrony, when both species were rare (stronger lower tail dependence) vs. both species were common (stronger upper tail dependence); (d) tail asymmetry for pairwise interaction matrix in a community computed as lower tail dependence minus upper tail dependence for synchronous species pair (I: independent, A: asynchronous, red cells: synchrony when species were rare, blue cells: synchrony when species were common); (e-h) four alternative model setups to test the effect of realms: community stability (inverse of variability) is the response, predictors are: richness(R), variance ratio(VR), total tail asymmetry(TA). Total tail asymmetry for a given community was defined as the sum of absolute values of red and blue cells from the tail asymmetry matrix shown in (d).
\label{fig_concept}}
\end{figure} 

\begin{figure}[!ht]
\includegraphics[width=18cm]{./Results/gather_res/simple_subset_birds_21/model_summary_resfig_ai2.pdf}
\caption{Results from four alternative model set-up (shown in Fig.1e-h): stability is different between two realms - freshwater communities were less stable than terrestrial ones. The relationships between stability and its predictors also varied between realms. The numbers shown are the medians of posterior distributions with 75$\%$ or 95$\%$ CI when significant indicated by stars. Non-significant estimates are shown by broken arrows and without stars.
 \label{fig_model_res}}
\end{figure} 

\begin{figure}[!ht]
\includegraphics[width=18cm]{./Results/gather_res/simple_subset_birds_21/plot_taxaeffect_ai.pdf}
\caption{Posterior estimates for the model including all predictors. Point estimates are median values (circles) with 75$\%$ (thicker bars), and 95$\%$ (thinner bars) credible intervals. The gray bars indicate the stability-difference between the realms (terrestrial; green minus freshwater; blue) are always significant from moderate to stronger scale. The hierarchical Bayesian model includes random intercepts for seven taxa, the effect size due to taxa is shown in (e).
 \label{fig_post_taxa}}
\end{figure} 

<!--Table 1: The comparison table among 8 possible models (4 without and 4 including realm), arranged in decreasing order of performance, shows the total tail asymmetry including realm was an important determinant than the usual richness and variance ratio predictors used to assess community stability. Abbreviations are ELPD = expected log pointwise predictive density; ELPD_SE = standard error in ELPD; LOOIC = leave-one-out information criterion; LOOIC_SE = standard error in LOOIC; WAIC = widely applicable information criterion; R2 = conditional R-squared; R2 (marg.) = marginal R-squared; RMSE = root-mean-square error. 
-->


<!--Appendix figure-->

\begin{figure}[!ht]
\includegraphics[width=18cm]{./Results/gather_res/datasummary_subset_birds_res/data_summary_map_ai_p3-pages-3.pdf}
\caption{Fig. A1: (a) Geographical locations for 1791 communities across two realms (freshwater: 716, terrestrial: 1075) and seven taxa (fish: 588, freshwater invertebrates: 113, freshwater plants: 15, birds: 293, terrestrial invertebrates: 159, terrestrial plants: 440, terrestrial mammals: 183). (b-c) two inclusion criteria: species-level abundance/ biomass data were collected for min. 20 years and species present for at least 70$\%$ of the study period, were considered for the analysis as species richness, R. \label{fig_data_summary_map}}
\end{figure} 

\begin{figure}[!ht]
\includegraphics[width=18cm]{./Results/gather_res/datasummary_subset_birds_res/data_summary_ai.pdf}
\caption{Boxplot diagram showing the distribution of the raw data values for response and predictor variables used in the model.
 \label{fig_data_summary}}
\end{figure} 

\begin{figure}[!ht]
\includegraphics[width=18cm]{./Results/gather_res/datasummary_subset_birds_res/syn_and_asyn_ratio_for_eachrealm.pdf}
\caption{Boxplot diagram showing the fraction of synchronous (pink) vs. asynchronous (green) interactions was much higher for the freshwater communities than the terrestrial ones.\label{fig_realm_ratio}}
\end{figure} 

